---
title: "Install required packages"
author: "David Kaiser"
date: "2020-04-18"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Description

This function checks if a specificed set of packages is installed and prompts the user for a decision to install or not. It is meant to be used within other functions that require packages to load.

## Arguments

+ *required* - names of required packages
+ *fun.name* = "Function" - A name for the function that requires the packages, defaults to "Function"

## Result

If any required packages are not installed the response Y (or y) will trigger the installation of missing packages. Any other response will call *stop()*, cancelling the function and the top function, and print a message containing the *fun.name*.

## Use within a function

```{r example within function, echo = TRUE, , eval=FALSE}
test.fun <- function(x){
  
  install_required_packages <- function(required,
                                        fun.name = "Function")
  {
    if(
      !all(required %in% installed.packages())
    ){
      missing <- required[!required %in% installed.packages()]
      response <- readline(prompt=cat("These required packages are missing:\n", missing, "\nDo you wish to install? [Y/N] "))
      if(response %in% c("y", "Y")){
        install.packages(missing)
        cat("\nRequired packages installed!\n\n")
      } else {
        stop(paste(fun.name, "interrupted. Required packages not installed. \n"), call. = FALSE)
      }
    }
  }
  
  install_required_packages(required = c("lubridate", "data.table", "ggplot2"), fun.name = "test.fun")
  
  return(x*10)
}
```

In this case I pretend that the function *test.fun* needs "lubridate", "data.table" and "ggplot2"; "lubridate" and "data.table" are not installed, "ggplot2" is. The output would look like this:

```{r eval=F}
test.fun(10)
These required packages are missing:
 lubridate data.table 
Do you wish to install? [Y/N] 
Y
Installing packages into [your_directory]

  There are binary versions available but the source versions are later:
           binary source needs_compilation
lubridate   1.7.4  1.7.8              TRUE
data.table 1.12.2 1.12.8              TRUE

  Binaries will be installed
trying URL 'https://cran.rstudio.com/bin/windows/contrib/3.4/lubridate_1.7.4.zip'
Content type 'application/zip' length 1367572 bytes (1.3 MB)
downloaded 1.3 MB

trying URL 'https://cran.rstudio.com/bin/windows/contrib/3.4/data.table_1.12.2.zip'
Content type 'application/zip' length 1926960 bytes (1.8 MB)
downloaded 1.8 MB

package ‘lubridate’ successfully unpacked and MD5 sums checked
package ‘data.table’ successfully unpacked and MD5 sums checked

The downloaded binary packages are in
	C:\Users\David-xps\AppData\Local\Temp\RtmpcjHvfW\downloaded_packages

Required packages installed!

[1] 100

```

Or like this if you chose not to install:
```{r eval=FALSE}
test.fun(10)
These required packages are missing:
 lubridate data.table 
Do you wish to install? [Y/N] 
N
Error: test.fun interrupted. Required packages not installed. 
```

